---
title: "COGS_score_contributions"
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
library(data.table)
library(ggplot2)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/COGS_results")

```

## Here is how to define genes that are specific to ILC3s, based on PIRs:
1) find the genes that are prioritised in hILC3s
2) Filter for the genes that require PIRs (or ABC enhancers) to be prioritised. This will be everything that is in the whole dataset that is not VProm+CodingOnly. 
3) find the genes that are prioritised in Javierre across all cell types.
4) Filter for those that require PIRs (i.e. everything not in VProm, coding again)
5) Make a table between the "ILC3 PIRs" genes and the "Javierre PIRs" genes.
6) Make another column of "ILC3 PIRs" genes that are also prioritised by VProm_coding only in Javierre. These are the ones prioritised by VProm because of HindIII.

#####
points 1 and 2.

```{r Get PIR genes in ILC3s  - this data is for frag res, 5kb and ABC interactions}
AI <- fread("./COGS_out/CD_deLange_ILCs_hg38_SuSIE_combinedInteractions_ALL/Annotated_COGS_scores_data.table.txt")
AV <- fread("./COGS_out/CD_deLange_ILCs_hg38_SuSIE_combinedInteractions_VProm_coding_only/Annotated_COGS_scores_data.table.txt") 

setnames(AI, "cogs", "cogs_total")
setnames(AV, "cogs", "cogs_VProm_coding")

AI_sig <- AI[cogs_total >= 0.5]
AV_bind <- AV[, .(ensg, cogs_VProm_coding)]

AI_AV <- AV_bind[AI_sig, on = "ensg"]
AI_AV[is.na(cogs_VProm_coding)] # none

PIRs_req <- AI_AV[cogs_VProm_coding < 0.5] 
PIRs_req[, ILC3_PIRs := "yes"]


```

####
Points 3 and 4.


```{r get PIR genes in Javierre}
J <- fread("./COGS_out/CD_deLange_Javierre_hg38_SuSIE_allCells/Annotated_COGS_scores_data.table_cell_type_summary_hg38.txt") # these are all the significant ones in Javierre in at least one cell type. Was re-run in hg38 using the hg38 rmap and baitmap.
JP <- fread("./COGS_out/CD_deLange_Javierre_hg38_SuSIE_allCells_VProm_coding_only/Annotated_COGS_scores_data.table.txt") # these are the ones that have been run on VProm Coding alone.

setnames(JP, "cogs", "cogs_VProm_coding")

JP_bind <- JP[, .(ensg, cogs_VProm_coding)]

J_JP <- JP_bind[J, on = "ensg"]
J_JP[is.na(cogs_VProm_coding)] 

PIRs_req_Jav <- J_JP[cogs_VProm_coding < 0.5] 
PIRs_req_Jav[, BLOOD_PIRs := "yes"]


```

points 5 and 6.

```{r merge ILC3s with Javierre}
PIRs_req_Jav_small <- PIRs_req_Jav[, .(ensg, BLOOD_PIRs)]

I_J1 <- PIRs_req_Jav_small[PIRs_req, on = "ensg"]

# Get the ones prioritised by VPromCoding in Javierre
JP_sig <- JP[cogs_VProm_coding >= 0.5]
JP_sig[, "BLOOD_VProm_Coding_alone" := "yes"] 
JP_sig_small <- JP_sig[, .(ensg, BLOOD_VProm_Coding_alone)]

I_J2 <- JP_sig_small[I_J1, on = "ensg"]
I_J2[is.na(I_J2)] <- "no"

final <- I_J2[, .(ensg, gene, cogs_total, ILC3_PIRs, BLOOD_PIRs, BLOOD_VProm_Coding_alone)]
setnames(final, c("cogs_total", "BLOOD_VProm_Coding_alone"), c("ILC3s_overall_COGS_score", "BLOOD_Promoter"))


```

Write the file. Get a col of prev prior evidence for these genes.

```{R write results}

prior <- fread("./COGS_out/CD_deLange_ILCs_hg38_SuSIE_combinedInteractions_ALL/Annotated_COGS_scores_data.table_with_prior_info_SuSIE_combinedInteractions.txt")
prior2 <- prior[, c("ensg", "Prior evidence")]
prior2[`Prior evidence` == "", prior := "no"]
prior2[`Prior evidence` != "", prior := "yes"]
prior2[, `Prior evidence` := NULL]
final_wprior <- final[prior2, on = "ensg", nomatch = NULL]

setnames(final_wprior, c("ILC3s_overall_COGS_score", "ILC3_PIRs", "BLOOD_PIRs", "BLOOD_Promoter", "prior"), c("ILC3s multiCOGS score", "Prioritised based on ILC3 PIRs", "Prioritised based on Javierre PIRs", "BLOOD_Promoter", "Prior mechanistic evidence in CD"))

out = "./comparisons/ILC_Javierre_PIR_analysis.txt"
fwrite(final_wprior, file = out, sep = "\t", quote = F, row.names = F, col.names = T)



```
